Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-05-10 ms/LMA_method.tex
remove ITV

\textbf{PREMISE:}
Leaf mass per area (LMA), which is the key trait in leaf economic spectrum and plant growth analysis, is measured from leaf discs or whole leaves.
These differences in the measurement methods may lead to large differences in the estimates of LMA values.

@@

Leaf mass per area (LMA), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area thus reflecting a cost for light interception, which is the key trait of the leaf economic spectrum (LES) (\protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Poorter2009}{Poorter et al., 2009}; \protect\hyperlink{ref-Onoda2017}{Onoda et al., 2017}).
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover (\protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}).
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns (\protect\hyperlink{ref-Garnier1994}{Garnier and Laurent, 1994}; \protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Reich2014}{Reich, 2014}; \protect\hyperlink{ref-Diaz2016}{Díaz et al., 2016}).
The LMA is also frequently used in plant growth analysis (\protect\hyperlink{ref-Evans1972}{Evans, 1972}; \protect\hyperlink{ref-Poorter2014}{Poorter et al., 2014}; \protect\hyperlink{ref-Falster2018}{Falster et al., 2018}) because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR \(\times\) LMR \(\times\) LMA\textsuperscript{-1}).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species (i.e., `soft' trait, \protect\hyperlink{ref-Diaz2004}{Díaz et al., 2004}).
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago (\protect\hyperlink{ref-Hanson1917}{Hanson, 1917}).

LMA values can be determined by either
1) a whole leaf including or excluding petioles, and
2) a leaf disc excluding all major veins or petioles (\protect\hyperlink{ref-Perez-Harguindeguy2013}{Pérez-Harguindeguy et al., 2013}; \protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}).
Although some studies do not make clear what protocol was followed, these differences in the measurement methods may lead to large discrepancies in the estimates of LMA values because the major vein allocation (major vein, which includes first to second or third-order veins, volume per area) has been reported to be one of the main determinants for the variation in LMA (\protect\hyperlink{ref-John2017}{John et al., 2017}).
Since larger leaves tend to invest more of their mass into dense midribs and petiole for support (\protect\hyperlink{ref-Niinemets2006}{Niinemets et al., 2006}, \protect\hyperlink{ref-Niinemets2007}{2007}), and thinner leaves have clearly visible and large-diameter veins with less uniform leaf structure (i.e., kite-type leaves; (\protect\hyperlink{ref-Grubb1986}{Grubb, 1986})), discrepancies in the estimates of whole-leaf LMA and leaf disc LMA might be greater for larger and thinner leaves.
If leaf disc LMA largely and consistently underestimates whole-leaf LMA, some calibrations may be required to combine or compare those different estimates (\protect\hyperlink{ref-Kraft2008}{Kraft et al., 2008}; \protect\hyperlink{ref-Onoda2011}{Onoda et al., 2011}).
If divergencies between the estimates based on the different methods are large and inconsistent (i.e., low \emph{R\textsuperscript{2}} values), this is difficult to calibrate and should inflate errors in the subsequent analyses.
To date, however, only a database from a single region, Panama Plant Traits Database (\protect\hyperlink{ref-Wright2010}{Wright et al., 2010}), is available in the TRY (\protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}) that have both estimates of LMA from small leaf discs and whole leaves including petioles, limiting of our understanding of under what conditions estimates of LMA from leaf discs are reliable estimates of whole-leaf LMA.

Given that leaf disc samples have more sources of intraspecific trait variation (ITV) than whole-leaf samples, the effects of discrepancies between whole-leaf LMA and leaf disc LMA might be large when ITV is quantified as a coefficient of variance (CV).
Sources of ITV for whole-leaf samples is variation among individuals within the same species and variation among leaves within the same individuals (\protect\hyperlink{ref-Messier2010}{Messier et al., 2010}, \protect\hyperlink{ref-Messier2017}{2017}).
An additional source of trait variation for leaf disc samples is variation among leaf discs within the same leaves.
Thus, extents of ITV for each species based on whole-leaf and leaf discs samples may not match.
Despite the importance of ITV in community ecology (\protect\hyperlink{ref-Kichenin2013}{Kichenin et al., 2013}; \protect\hyperlink{ref-Siefert2015}{Siefert et al., 2015}), the effect of different measurement methods on the extent of ITV has been largely ignored.

In this study, we aimed to investigate what extent whole-leaf (including petiole) and leaf disc LMA match, whether the relationships between whole-leaf LMA and leaf disc LMA varied in leaf sizes and leaf thicknesses, and whether amounts of intraspecifc variations for each match between whole-leaf and leaf disc based estimates.
We collected a total of 334 woody species from four biomes (tropics, subtropics, savanna, and warm-temperate) to cover the wide range of geography.
We predicted that the overall relationship between whole-leaf LMA and disc LMA would be weaker than previous reported LMA data (\emph{R\textsuperscript{2}} = 0.92 for whole-leaves in Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} = 0.92 for leaf laminas in Onoda et al. (\protect\hyperlink{ref-Onoda2011}{2011})) because we collected data with various leaf sizes from a wide range of climates.
We also predicted that species with larger and/or thinner leaves would show weaker correlations.
Finally, we predicted that amounts of intraspecifc variations for each species do not match well between whole-leaf and leaf disc based estimates.

@@

We used newly compiled individual-level plant datasets form a wide ranges of biomes in China and Japan (Table S1) to examine the relationship between whole-leaf and leaf disc-based estimates of leaf traits.
First, the Yunnan dataset is from three forest plots in Yunnan province, Southwest China, which includes a tropical rain forest (TRF), a hot-dry savanna ecosystem (HDS), and a subtropical evergreen wet forest (STF), with a total of species and individuals.
Second, the Yakushima data is from a warm-temperate forest in Yakushima island, Japan, with a total of 193 species and individuals.
In total, we used 334 woody species and individuals.
The methodologies for trait sampling and measurement are slightly different between the Yunnan and the Yakushima dataset, which we describe bellow.

@@

@@

To determine what extent LMA values were affected by the different methods, LMA were determined (i) LMA based on a whole leaf including petioles and (ii) LMA based on leaf disc avoiding thick veins.
Three 0.6-cm diameter discs were taken from the base, middle and tip of each leaf using a hole punch in the Yunnan dataset.
Two 1.0-cm diameter discs were taken from each leaf were in the Yakushima dataset.
Fresh leaf area (LA; cm\textsuperscript{-2}) of leaf materials (whole leaves including petioles and midribs) was measured using a scanner and ImageJ software by the R package \emph{LeafArea} (\protect\hyperlink{ref-Katabuchi2015}{Katabuchi, 2015}).
Fresh leaf thickness (LT; mm) was measured using a micrometer in three points at the base, middle and tip of the leaf using a micrometer (Mitutoyo 293-240, Mitutoyo Corporation, Kawasaki, Japan) with a precision at 0.001 mm.
Thickness of each leaf disc was also measured for the Yunnan dataset but not for the Yakushima dataset, which enables us to estimate whole-leaf and leaf disc leaf tissue density (LD; g cm\textsuperscript{-3}) separately in the Yunnan dataset.
Leaf dry mass (whole leaves including petioles and midribs, and leaf discs) was recorded after drying 80°C for over 48 h to constant weight using an electronic scale (Mettler-Toledo MS204TS, Columbus, OH, U.S.A.) with a precision at 0.0001 g.
In the subset of Yakushima dataset, leaf lamina and petiole dry mass were also determined separately (86 out of 193 species) to quantify leaf support costs.
Dry mass for leaf discs were recorded 2-3 discs together for each leaf.
Based on those measurements of whole-leaf and leaf disc samples, LMA (g m\textsuperscript{-2}) was calculated as the ratio of leaf dry mass to fresh leaf area, and LD was calculated as the ratio of LMA to LT.
We do not present results of LD in the main text, because differences between whole-leaf LD and leaf disc LD only depends on the ratio between whole-leaf LMA and leaf disc LMA and measurement errors of leaf thickness (Appendix S2).

@@

\emph{Bivariate trait relationships for mean and CV}

Bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were further summarized with variance explained (\emph{R\textsuperscript{2}}) based on Pearson correlations and with standardized major axis (SMA) regressions (\protect\hyperlink{ref-Warton2006}{Warton et al., 2006}) across all the samples and different leaf area and thickness groups.

Species groups were divided into four categories based on the medians of leaf area and leaf thickness across all species:
thin and small-leaved, thin and large-leaved, thick and small-leaved, thick and large-leaved species.
Median values are as following:
small-leaved species (LA \textless{} cm\textsuperscript{2}),
and thin-leaved species (LT \textless{} mm).

The amount of intraspecific variation (ITV) was calculated using Bao's CV estimator (\protect\hyperlink{ref-Yang2020}{Yang et al., 2020});

\begin{align}
CV = CV_1 - \frac{CV_1^3}{N} + \frac{CV_1}{4N} + \frac{CV_1^2 \gamma_1}{2N} + \frac{CV_1 \gamma_2}{8N} \label{eq:CV}
\end{align}

where \(CV_1 = s/\bar{x}\), \(s\) is the sample standard deviation and \(\bar{x}\) is the sample mean, \emph{N} is the sample size, and \(\gamma_1\) and \(\gamma_2\) are the skewness and kurtosis of the trait distribution.
Bao's CV estimator is a robust method to quantify ITV based on coefficient of variation (CV) when sample size is small (\protect\hyperlink{ref-Yang2020}{Yang et al., 2020}).

SMA analyses were run using the R package \emph{smatr} (\protect\hyperlink{ref-Warton2012a}{Warton et al., 2012}); this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.
All statistical analyses were conducted in R version 4.1.3 (\protect\hyperlink{ref-RCoreTeam2022}{R Core Team, 2022}).

@@

There were a significant shift in elevation between whole-leaf and leaf disc LMA (Fig.1, SMA slope: 0.975 {[}95\%CI: 0.938, 1.013{]}, SMA intercept: 0.0881 {[}95\%CI: 0.016, 0.16{]}), suggesting that the whole-leaf estimates of LMA were 9.69\% {[}95\%CI: 8.07, 11.4{]} greater than disc-based estimates in the arithmetic scale.

Leaf tissue density, leaf area, leaf thickness and leaf punch size influenced the relationship between the whole-leaf LMA and leaf disc LMA (Fig. 2).
When a large leaf punch was used, species with lower leaf tissue density had grater variance in the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2a).
Species with lower leaf tissue density had greater estimate whole-leaf to leaf disc LMA ratio when a small leaf punch was used (Fig. 2a).
Species with larger leaves had greater estimate of the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2b), which was only significant for leaves obtained by a large punch.
Species with thinner leaves had grater variance in the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2c), which was only significant for leaves obtained by a small punch.
Posteriors of coefficients are summarized in (Fig. SX).
Leaf disc and whole-leaf LMA estimate diverged for species with greater investment in petiole dry mass (Fig. 3a). Large-leaved species tended to have greater investment for petiole dry mass than small-leaved species (Fig. 3b).

Amounts of intraspecific variation (CV; coefficient of variation) only weakly matched between whole-leaf and disc-based estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.33).

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Kichenin2013}{}}%
Kichenin, E., D. A. Wardle, D. A. Peltzer, C. W. Morse, and G. T. Freschet. 2013. \href{https://doi.org/10.1111/1365-2435.12116}{Contrasting effects of plant inter- and intraspecific variation on community-level trait measures along an environmental gradient}. \emph{Functional Ecology} 27: 1254--1261.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Messier2017}{}}%
Messier, J., B. J. McGill, B. J. Enquist, and M. J. Lechowicz. 2017. \href{https://doi.org/10.1111/ecog.02006}{Trait variation and integration across scales: Is the leaf economic spectrum present at local scales?} \emph{Ecography} 40: 685--697.

\leavevmode\vadjust pre{\hypertarget{ref-Messier2010}{}}%
Messier, J., B. J. McGill, and M. J. Lechowicz. 2010. \href{https://doi.org/10.1111/j.1461-0248.2010.01476.x}{How do traits vary across ecological scales? {A} case for trait-based ecology}. \emph{Ecology Letters} 13: 838--848.

@@

\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam2022}{}}%
R Core Team. 2022. R: {A} language and environment for statistical computing. {R Foundation for Statistical Computing}, {Vienna, Austria}.

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yang2020}{}}%
Yang, J., J. Lu, Y. Chen, E. Yan, J. Hu, X. Wang, and G. Shen. 2020. Large {Underestimation} of {Intraspecific Trait Variation} and {Its Improvements}. \emph{Frontiers in Plant Science} 11.

@@

\textbf{Fig. 4.}
Relationships between coefficient of variation (Bao's CV estimator) in leaf trait values determined by using whole leaves and leaf discs: (a) leaf mass per area (LMA).
The correlation is significant (\emph{P} \textless{} 0.001).
Details as in Fig. 1.
Note that the axes are square root scales.

>>>>

\textbf{PREMISE:}
Leaf mass per area (LMA), which is an important functional trait in leaf economic spectrum and plant growth analysis, is measured from leaf discs or whole leaves.
These differences in the measurement methods may lead to large differences in the estimates of LMA values.

@@

The primary function of leaves is to return photosynthetic revenue on the investment that has been made in constructing the leaf (i.e., leaf economic spectrum, \protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Westoby2013}{Westoby et al., 2013}).
Leaf mass per area (LMA), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area, which is the key trait of the leaf economic spectrum (\protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Poorter2009}{Poorter et al., 2009}; \protect\hyperlink{ref-Onoda2017}{Onoda et al., 2017}).
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover (\protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}).
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns (\protect\hyperlink{ref-Garnier1994}{Garnier and Laurent, 1994}; \protect\hyperlink{ref-Wright2004a}{Wright et al., 2004}; \protect\hyperlink{ref-Reich2014}{Reich, 2014}; \protect\hyperlink{ref-Diaz2016}{Díaz et al., 2016}).
The LMA is also frequently used in plant growth analysis (\protect\hyperlink{ref-Evans1972}{Evans, 1972}; \protect\hyperlink{ref-Poorter2014}{Poorter et al., 2014}; \protect\hyperlink{ref-Falster2018}{Falster et al., 2018}) because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR \(\times\) LMR \(\times\) LMA\textsuperscript{-1}).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species (i.e., `soft' trait, \protect\hyperlink{ref-Diaz2004}{Díaz et al., 2004}).
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago (\protect\hyperlink{ref-Hanson1917}{Hanson, 1917}).
Actually, LMA was the most often requested leaf trait in the TRY plant data database (2,977 out of 7,330 requests), followed by leaf nitrogen contents per leaf dry mass (1,938 requests, \protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020} {[}status October 1 2019{]}).

LMA values can be determined by either
1) a whole leaf including (88,490 observations and 7,068 species) or excluding petioles (64,838 observations and 7,558 species; note that this could include leaf disc samples), and
2) a leaf disc excluding all major veins or petioles (645 observations and 403 species) (\protect\hyperlink{ref-Perez-Harguindeguy2013}{Pérez-Harguindeguy et al., 2013}; \protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020} {[}TRY version 5{]}).
Although many studies do not make clear what protocol was followed (146,315 observations and 13,101 species), these differences in the measurement methods may lead to large discrepancies in the estimates of LMA values because the major vein allocation (major vein, which includes first to second or third-order veins, volume per area) has been reported to be one of the main determinants for the variation in LMA (\protect\hyperlink{ref-John2017}{John et al., 2017}).
Previous works that compared estimates of whole-leaf LMA and leaf disc LMA using tropical tree species showed good correlations between the two Onoda et al. (\protect\hyperlink{ref-Onoda2011}{2011}), but the comparisons between the two estimates in other biomes have rarely been examined.

Since larger leaves tend to invest more of their mass into dense midribs and petiole for support (\protect\hyperlink{ref-Niinemets2006}{Niinemets et al., 2006}, \protect\hyperlink{ref-Niinemets2007}{2007}), and thinner leaves have clearly visible and large-diameter veins with less uniform leaf structure (i.e., kite-type leaves; (\protect\hyperlink{ref-Grubb1986}{Grubb, 1986})), discrepancies in the estimates of whole-leaf LMA and leaf disc LMA might be greater for larger and thinner leaves.
(\textbf{add leaf density})
If leaf disc LMA largely and consistently underestimates whole-leaf LMA (i.e., differences in means), some calibrations may be required to combine or compare those different estimates (\protect\hyperlink{ref-Kraft2008}{Kraft et al., 2008}; \protect\hyperlink{ref-Onoda2011}{Onoda et al., 2011}).
If divergencies between the estimates based on the different methods are large and inconsistent (i.e., differences in variances and low \emph{R\textsuperscript{2}} values), this is difficult to calibrate and should inflate errors in the subsequent analyses.
To date, however, only a database from a single region, Panama Plant Traits Database (\protect\hyperlink{ref-Wright2010}{Wright et al., 2010}), is available in the TRY (\protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}) that have both estimates of LMA from small leaf discs and whole leaves including petioles, limiting of our understanding of under what conditions estimates of LMA from leaf discs are reliable estimates of whole-leaf LMA.

In this study, we aimed to investigate what extent whole-leaf (including petiole) and leaf disc LMA match, whether the relationships between whole-leaf LMA and leaf disc LMA varied in leaf tissue density, leaf area and leaf thickness, and whether amounts of intraspecifc variations for each match between whole-leaf and leaf disc based estimates.
We collected a total of 334 woody species from four biomes (tropics, subtropics, savanna, and warm-temperate) to cover the wide range of geography.
We predicted that the overall relationship between whole-leaf LMA and disc LMA would be weaker than previous reported LMA data (\emph{R\textsuperscript{2}} = 0.92 for whole-leaves in Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} = 0.92 for leaf laminas in Onoda et al. (\protect\hyperlink{ref-Onoda2011}{2011})) because we collected data with various leaf sizes from a wide range of climates.
We also predicted that species with larger and/or thinner leaves would show weaker correlations.
Finally, we predicted that amounts of intraspecifc variations for each species do not match well between whole-leaf and leaf disc based estimates.

@@

We used newly compiled individual-level plant datasets form a wide ranges of biomes in China and Japan (Table S1) to examine the relationship between whole-leaf and leaf disc-based estimates of leaf traits.
First, the Yunnan dataset is from three forest plots in Yunnan province, Southwest China, which includes a tropical rain forest (TRF), a tropical hot-dry savanna ecosystem (HDS), and a subtropical evergreen wet forest (STF), with a total of species and individuals.
Second, the Yakushima data is from a warm-temperate forest in Yakushima island, Japan, with a total of 193 species and individuals.
In total, our data set comprised 334 woody species and 1365 individual that ranged from xxx to 24.7 °C in mean annual temperature (MAT) and 732.8 to 4477 mm in mean annual precipitation (MAP) {[}Table S1; Eguchi (\protect\hyperlink{ref-Eguchi2006}{2006}); Song et al. (\protect\hyperlink{ref-Song2017}{2017}); Fei et al. (\protect\hyperlink{ref-Fei2018}{2018}){]}, which covers the wide climate ranges in the distribution of broad-leaved tree species.
The methodologies for trait sampling and measurement are slightly different between the Yunnan and the Yakushima dataset, which we describe bellow.

@@

We excluded species that had tiny leaves or leaflets (length \textless1 cm) because it was not practical to measure leaf disc LMA.

@@

To determine what extent LMA values were affected by the different methods, LMA were determined (i) LMA based on a whole leaf including petioles and (ii) LMA based on leaf disc avoiding thick veins.
Three 0.6-cm diameter discs were taken from the base, middle and tip of each leaf using a hole punch in the Yunnan dataset.
Two 1.0-cm diameter discs were taken from each leaf were in the Yakushima dataset.
Fresh leaf area (LA; cm\textsuperscript{-2}) of leaf materials (whole leaves including petioles and midribs) was measured using a scanner and ImageJ software by the R package \emph{LeafArea} (\protect\hyperlink{ref-Katabuchi2015}{Katabuchi, 2015}).
Fresh leaf thickness (LT; mm) was measured using a micrometer in three points at the base, middle and tip of the leaf using a micrometer (Mitutoyo 293-240, Mitutoyo Corporation, Kawasaki, Japan) with a precision at 0.001 mm.
Thickness of each leaf disc was also measured for the Yunnan dataset but not for the Yakushima dataset, which enables us to estimate whole-leaf and leaf disc leaf tissue density (LD; g cm\textsuperscript{-3}) separately in the Yunnan dataset.
Leaf dry mass (whole leaves including petioles and midribs, and leaf discs) was recorded after drying 80°C for over 48 h to constant weight using an electronic scale (Mettler-Toledo MS204TS, Columbus, OH, U.S.A.) with a precision at 0.0001 g.
In the subset of Yakushima dataset, leaf lamina and petiole dry mass were also determined separately (80 out of 193 species) to quantify leaf support costs.
Dry mass for leaf discs were recorded 2-3 discs together for each leaf.
Based on those measurements of whole-leaf and leaf disc samples, LMA (g m\textsuperscript{-2}) was calculated as the ratio of leaf dry mass to fresh leaf area, and LD was calculated as the ratio of LMA to LT.
We do not present results of LD in the main text, because differences between whole-leaf LD and leaf disc LD only depends on the ratio between whole-leaf LMA and leaf disc LMA and measurement errors of leaf thickness (Appendix S2).

\emph{Relationship between whole-leaf and leaf disc LMA}

A bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were summarized with variance explained (\emph{R\textsuperscript{2}}) based on Pearson correlations and with standardized major axis (SMA) regressions (\protect\hyperlink{ref-Warton2006}{Warton et al., 2006}) across all the species.

SMA analyses were run using the R package \emph{smatr} (\protect\hyperlink{ref-Warton2012a}{Warton et al., 2012}); this included tests for slope heterogeneity and elevation differences between disc-based and whole-leaf estimates of LMA values.
Confidence intervals of SMA estimates were based on 2000 bootstraps.

@@

We also perfomred ordinary least-squares (OLS) regressions to see the relationship between whole-leaf to leaf disc LMA ratio and petiole to leaf dry mass ratio and between leaf area petiole to leaf dry mass ratio, using a subset of the dataset in which we measured leaf lamina and petiole dry mass separately.
All statistical analyses were conducted in R version 4.1.3 (\protect\hyperlink{ref-RCoreTeam2021}{\textbf{RCoreTeam2021?}}).

@@

There were a significant shift in elevation between whole-leaf and leaf disc LMA (Fig.1, SMA slope: 0.975 {[}95\%CI: 0.938, 1.013{]}, SMA intercept: 0.0881 {[}95\%CI: 0.016, 0.16{]}), suggesting that the whole-leaf estimates of LMA were 9.68\% {[}95\%CI: 8.05, 11.4{]} greater than disc-based estimates in the arithmetic scale.
Leaf tissue density, leaf area, leaf thickness and leaf punch size influenced the relationship between the whole-leaf LMA and leaf disc LMA (Fig. 2).
When a large leaf punch was used, species with lower leaf tissue density had grater variance in the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2a).
Species with lower leaf tissue density had greater estimate whole-leaf to leaf disc LMA ratio when a small leaf punch was used (Fig. 2a).
Species with larger leaves had greater estimate of the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2b), which was only significant for leaves obtained by a large punch.
Species with thinner leaves had grater variance in the estimate of whole-leaf to leaf disc LMA ratio (Fig. 2c), which was only significant for leaves obtained by a small punch.
Posteriors of coefficients are summarized in (Fig. SX).
Leaf disc and whole-leaf LMA estimate diverged for species with greater investment in petiole dry mass (Fig. 3a). Large-leaved species tended to have greater investment for petiole dry mass than small-leaved species (Fig. 3b).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Eguchi2006}{}}%
Eguchi, T. 2006. Climate. Yakushima, 3--26. {Asakura Shoten}, {Tokyo (in Japanese)}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Fei2018}{}}%
Fei, X., Q. Song, Y. Zhang, Y. Liu, L. Sha, G. Yu, L. Zhang, et al. 2018. \href{https://doi.org/10.1016/j.scitotenv.2017.10.239}{Carbon exchanges and their responses to temperature and precipitation in forest ecosystems in {Yunnan}, {Southwest China}}. \emph{Science of The Total Environment} 616--617: 824--840.

@@

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Song2017}{}}%
Song, Q.-H., Y.-P. Zhang, L.-Q. Sha, X.-B. Deng, Y. Deng, C.-S. Wu, Z.-Y. Lu, et al. 2017. \href{https://doi.org/10.3832/ifor2223-010}{Canopy temperature variability in a tropical rainforest, subtropical evergreen forest, and savanna forest in {Southwest China}}. \emph{iForest - Biogeosciences and Forestry} 10: 611.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Westoby2013}{}}%
Westoby, M., P. B. Reich, and I. J. Wright. 2013. \href{https://doi.org/10.1111/nph.12345}{Understanding ecological variation across species: {Area-based} vs mass-based expression of leaf traits}. \emph{New Phytologist} 199: 322--323.

@@

@@

\textbf{Fig. 4.}
Relationships between coefficient of variation (Bao's CV estimator, Eq. 5) in leaf trait values determined by using whole leaves and leaf discs: (a) leaf mass per area (LMA).
The correlation is significant (\emph{P} \textless{} 0.001).
Details as in Fig. 1.
Note that the axes are square root scales.
